Viscous Transport in Eroding Porous Media

### 2019-09-26 docs/transport/erosion_transport.tex
edit in transport paper
working on section 6

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a): The distribution of the pore
sizes of an eroded geometry. The black curve is the Weibull distribution
with the same first two moments as the data. (b): We only consider pore
sizes between neighboring grains, where two grains are neighbors if they
share a edge in a Delaunay triangulation using nodes placed at the
center of each eroding grain.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo[inline]{here}
In this section, we apply our new quadrature method to (1) simulate
erosion of several different configurations of eroding bodies, and (2)
characterize transport inside eroded geometries.  Each eroding body is
initially a circle with center $\cc_i$ and radius $r_i$.  The center and
radius are chosen at random, and the body is accepted as long as it is
contained in $[-1,1] \times [-1,1]$ and is sufficiently separated from
all other bodies.  Since the error of our quadrature method is bounded
independent of the geometry, we consider bodies that are separated by
less than 10\% of an arclength spacing.  The randomized method is
repeated until the initial geometry reaches a desired initial porosity.

Once the geometry is initialized, each body has length $L_i = 2\pi r_i$,
and they are each discretized with $N_\iin$ points.  A no-slip boundary
condition is imposed on each eroding body $\gamma_i$.  For all
simulations, we discretize the outer wall $\Gamma$ with $N_\out=1024$
points.  For all but the first example, the Dirichlet boundary condition
on $\Gamma$ is a hagen-poiseuille flow.  As the bodies erode, the flow
rate is adjusted to maintain the physical constant pressure drop
boundary condition.  To maintain this pressure drop, we compute the
pressure at near the inlet and outlet at each time step as described in
our previous work~\cite{qua-moo2018}.

In each of the examples, we include plots of the vorticity in the fluid
bulk.  Since the vorticity and the shear stress agree on surface of the
eroding bodies, the vorticity is a useful method to visualize the
erosion rate.  In particular, the rate of erosion is largest in regions
where the magnitude of the vorticity is largest.

Because of the definition of the erosion rate, there is a loss of
regularity, and inevitably leads to corners developing on the
eroding bodies.  Therefore, as described in
Section~\ref{sec:formulation} and our previous work~\cite{qua-moo2018},
we introduce a curvature penalization term and a Gaussian smoothing
step.  For all examples, we use the smoothing parameters $\eps=15/256$
and $\sigma=10/256$.

The common characteristic of each of the experiments is near-contact
between the eroding bodies, outer walls, and streamlines. We use our
numerical methods to describe and visualize the following examples:
\begin{itemize}
  \item{\bf Single Body Close to a Wall}: We consider a single eroding
  body close to the outer wall at $y=-1$.  Using a shear flow that
  vanishes on $y=-1$, we compare the eroding body's shape to a similar
  experiment of Spagnolie and Mitchell~\cite{mit-spa2017}.

@@

  \item{\bf 100 Bodies at a Medium Porosity}: We consider 100 eroding
  bodies with a medium porosity.  We compute the tortuosity, anomalous
  dispersion rates, and the pore throat distributions.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a single eroding body close to a solid wall.  To compare our
results with a similar numerical experiment performed by Mitchell and
Spagnolie~\cite{mit-spa2017}, rather than a pressure drop boundary
condition, we impose the shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
on the outer wall.  In their experiment, Mitchell and Spagnolie place
the center of the eroding body $1.5$ times the body's radius above the
solid wall.  By using the Barycentric quadrature rule described in
Section~\ref{sec:bary} rather than the trapezoid rule, the eroding body
can be placed much closer to the solid wall without requiring an
unreasonable number of discretization points.  We consider an eroding
body that is initially a circle of radius $r=0.4$.  We discretize the
eroding body with $N_\iin=256$ discretization points, and its center is
chosen so that the distance between the body and the solid wall is
$h/2$, $h/4$, and $h/10$, where $h = 2\pi r/N_\iin$.  If we were to use
the trapezoid rule and require an error that is comparable to the
Barycentric quadrature rule, the body with an initial distance of $h/10$
from the solid wall would require $6,400$ discretization points, and the
outer wall would require $50,000$ discretization points.  For all three
simulations, we use the time step size $\Delta t=10^{-4}$.

In Figure~\ref{fig:NearWall}, we superimpose the eroding body's shape at
equispaced time steps.  For the three initial configurations, the shear
stress is positive for all time, but varies over several orders of
magnitude.  Therefore, we color the eroding body's boundary with the
logarithm of the shear stress.  Since the shear stress is always
positive, the erosion rate is smooth and corners do not develop.
However, the shear stress varies over several orders of magnitude, and
this leads to interesting shapes.  For instance, in the top half of the
body, there is a sudden increase in the shear stress, and this leads to
a region of high curvature.  This behavior is also present in three
dimensions~\cite{mit-spa2017} (Figure 7(c)).  The biggest difference
between the two- and three-dimensional results is the presence of a
recirculation zone.  In three dimensions, there is no recirculation
between the solid wall and the spherical body~\cite{cha-feu2003}.  In
contrast, the two-dimensional flow Stokes flow can have recirculation
zones~\cite{chw-wu1975, hig1985}. To visualize the flow, we plot the
vorticity of the final time step from Figure~\ref{fig:NearWall} in
Figure~\ref{fig:NearWall_vort}.  The recirculation is present in the
region where the vorticity is smallest.

\begin{figure}[htbp]
\begin{center}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r1h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d5h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.38\textwidth}
\includegraphics[width =\textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{}
\end{subfigure}
\caption{\label{fig:NearWall} A single body eroding in a Stokes flow.
The color is the logarithm of the shear stress. Therefore, erosion is
fastest in the red regions (upper half) and slowest in the blue regions
(lower half).  The body is initialized at three different distances from
the lower wall: (a) $d=h$, (b) $h/2$, and (c) $h/10$.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r1h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d5h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d1h_vort}
\caption{}
\end{subfigure}
\caption{\label{fig:NearWall_vort} The vorticity of the fluid with a
single body eroding at time t=0.1. The initial distance from the body to
the solid wall are: (a) $h$, (b) $h/2$, and (c) $h/10$.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points and a time step size of $\Delta t = 10^{-4}$.  The background
flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 2.  Snapshots of the bodies at four equispaced
times are shown in Figure~\ref{fig:Eroding20vort}.  The color is the
vorticity which is equivalent to the shear stress when restricted to the
eroding bodies.  Therefore, the rate of erosion is fastest in regions
where the magnitude of the vorticity is largest.  Initially, several of
the eroding bodies are closer to the outer wall than the $5h$ threshold
required to guarantee that the trapezoid rule achieves machine
precision.  In particular, with $h=L_2/N_\out$, the distance between
bodies 1, 6, 13, and 15 and the outer wall are $1.3h$, $2.9h$, $2.8h$,
and $1.3h$, respectively.  In addition, the distance between several
pairs of eroding bodies is too small to be accurately resolved with the
trapezoid rule at the prescribed resolution.  Particular pairs include
bodies 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18.  By using the Barycentric
quadrature rule, the interaction between these nearly-touching bodies
can be accurately resolved at the prescribed resolution, and erosion can
be simulated until all the bodies have vanished.

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} Erosion of 20 nearly touching bodies
with the fixed pressure drop condition. The color is the vorticity of
the fluid. The four snapshots are evenly spaced in time. We use $N_\iin
= 256$ discretization points on the bodies.  In the fourth frame, bodies
4, 12, and 17 have vanished completely, and body 19 has almost
completely eroded.}
\end{center}
\end{figure}

We observe that erosion creates a network of channels from the inlet to
outlet where the velocity and vorticity, and therefore erosion rate, are
much larger when compared to other regions.  These channels can be
further visualized by considering the streamlines.  In
Figure~\ref{fig:Eroding20tracer}, we freeze the geometry at the second
time step from Figure~\ref{fig:Eroding20vort}, and we plot 200
streamlines that are initialized at $x=-1$ and equispaced in $y$.  The
streamlines are shown at five different times, and the final plot is a
zoom in of the lower right quadrant of the fifth time step.  Since we
use a high-order quadrature rule and a high-order time stepping method,
we resolve streamlines that come very close to the eroded bodies.  There
are three clear regions where the streamlines are concentrated, and this
corresponds to regions of high velocity.  Two of these regions are
located between the bodies and the solid walls at $y=\pm 1$, and the
third cuts through the porous region with the upper part of the channel
formed by bodies 1, 4, 6, and 8.  Since the flow is fastest in these
regions, the shearing is largest, and this causes these openings to grow
fastest which can be observed in Figure~\ref{fig:Eroding20vort}.

Similar to our previous work, erosion causes the space between nearly
touching bodies to quickly expand, and flat faces develop along the
region of near contact.  This qualitative behavior is present in
Figure~\ref{fig:Eroding20vort} between bodies 3 \& 4, 15 \& 16, and
others.  However, by resolving the interaction betweeen bodies that are
much closer together than in our previous work, we observe that, at
least initially, very little erosion occurs between certain pairs of
bodies.  For instance the opening between bodies 1 \& 6, 3 \& 9, and 5
\& 13 grow much slower than the opening between bodies 15 \& 16.  A
common feature of the openings that grow slowly is that they are nearly
perpendicular to the main flow direction, and this results in a small
velocity, and hence a slow erosion rate.  In fact, when plotting the
streamlines, only a small fraction of the plotted streamlines pass
between theese pairs of bodies indicating a very small velocity.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from Figure~\ref{fig:Eroding20vort}. The streamlines are
initialized at $x=-1$ and are equally spaced in the $y$-direction. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot.  Since we use the Barycentric
quadrature rule and the fourth-order Runge-Kutta method, the streamlines
can be very close to but not unphysically penetrate the bodies.}
\end{center}
\end{figure}

Next, we use the streamlines to compute the local tortuosity and the
anomalous dispersion rate of eroding geometries.  Since these
calculations are statistical, we use $N_p = 1000$
streamlines~\cite{bel-sal-rin1992}.  To compute the tortuosity, we
require the velocity at the start of the streamlines.  These normalized
velocities are plotted in Figure~\ref{fig:Eroding20tort}(a) for the
eroded geometry with porosity 62.9\% (see
Figure~\ref{fig:Eroding20tort}(c)).  The velocities are similar to the
work of Matyka et al.~\cite{mat-kha-koz2008} (see Figure 4(a)), except
that our cross-section, by construction, does not cut through any of the
bodies.  Next, in Figure~\ref{fig:Eroding20tort}(b), we plot the local
tortuosity~\eqref{eqn:localTort} by calculating the relative length of
each streamline as it traverses the channel from $x=-1$ to $x=1$.  The
local tortuosity ranges from 1 to 1.27, meaning that one of the
streamlines is 27\% longer than it would have been if the bodies were
absent.  The average streamline length is 9.79\% longer than if the
bodies had been absent, so the tortuosity of the eroded channel is
$1.098$.  Again, comparing the local tortuosity to
Matyka~\cite{mat-kha-koz2008} (Figure 4(b)), the results are
qualitatively similar. However, since our initial cross-section does not
cut through any bodies, the local tortuosity does not have any gaps.
The discontinuities in the local tortuosity happens neighboring
streamlines diverge on opposite sides of a body.  In
Figure~\ref{fig:Eroding20tort}(c), we plot pairs of streamline that
correspond to the ten largest jumps in the local tortuosity.
Corresponding pairs are plotted in the same color.

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =0.97\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity of an eroded geometry
with porosity 62.9\%. (a) The $x$-component of the velocity, $u(-1,y)$,
normalized by its maximum velocity of $2.98 \times 10^{-3}$. (b) The
local tortuosity $\tau(y)$ along the cross section $x = -1$.  The
spatially averaged tortuosity through the channel is $1.098$.  (c) The
streamlines resulting in the ten largest differences between neighboring
streamlines.  Neighboring streamlines have the same color.}
\end{figure}

In Figure~\ref{fig:Eroding20Transport}(a), we plot the tortuosity as a
function of the porosity. The initial porosity is $\phi = 37.68\%$, and
the initial tortuosity is $1.16$.  The tortuosity is computed with both
the length of the streamlines~\eqref{eqn:tortuosity1} (red marks) and
using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in Figure~\ref{fig:Eroding20tort}(c).
The two tortuosity formulas give similar results, and any discrepancy
can be accounted for by slow regions of recirculation and quadrature
error induced by the integrals in the tortuosity definitions.  As the
bodies erode, this creates wide channels where streamlines undergo only
minor vertical deflections, and this explains why the tortuosity
eventually decreases with porosity.  We computed lines of best fit using
the porosity-tortuosity models~\eqref{eqn:tortuosityModels}, and the
power law model gave the smallest error.  The  black dashed line in
Figure~\ref{fig:Eroding20Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064}
\end{align}
with a root-mean-square error of $5.90 \times 10^{-3}$.  The initial
increase in tortuosity occurs because in the absence of erosion (left
plot in Figure~\ref{fig:Eroding20vort}), many of the streamlines, such
as those initialized between bodies 15 \& 18, only perform minor
deflections to pass through the narrow regions, albeit, very slowly.
However, as erosion starts to open the channels, the streamlines deflect
into the fast regions, such as the region above body 11, and this
increases the amount of vertical deflection, and therefore the
tortuosity.

>>>>

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a) The distribution of the pore
sizes of an eroded geometry. The black curve is the Weibull distribution
with the same first two moments as the data. (b) We only consider pore
sizes between neighboring grains, where two grains are neighbors if they
share a edge in a Delaunay triangulation using nodes placed at the
center of each eroding grain.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this section, we apply our new quadrature method to simulate
erosion of several different configurations of nearly-touching eroding
bodies, and to characterize transport inside eroded geometries.  Each
eroding body is initially a circle with center $\cc_i$ and radius $r_i$.
The center and radius are chosen at random, and the body is accepted as
long as it is contained in $[-1,1] \times [-1,1]$ and is sufficiently
separated from all other bodies.  Since the error of our quadrature
method is bounded independent of the geometry, we consider bodies that
are separated by less than 10\% of an arclength spacing.  The randomized
method is repeated until the initial geometry reaches a desired initial
porosity.

Once the geometry is initialized, each body has length $L_i = 2\pi r_i$,
and they are each discretized with $N_\iin$ points.  A no-slip boundary
condition is imposed on each eroding body $\gamma_i$.  For all
simulations, we discretize the outer wall $\Gamma$ with $N_\out=1024$
points.  For all but the first example, the Dirichlet boundary condition
on $\Gamma$ is a Hagen-Poiseuille flow, and the flow rate is adjusted at
each time step to maintain a constant pressure drop. Since the fluid
equations are linear, this is achieved by computing the pressure near
the inlet and outlet at each time step, and then scaling the flow rate
appropriately~\cite{qua-moo2018}.

In each of the examples, we plot the vorticity in the fluid bulk.  Since
the vorticity and the shear stress agree on surface of the eroding
bodies, the vorticity is a useful method to visualize the erosion rate.
In particular, the rate of erosion is largest in regions where the
magnitude of the vorticity is largest.

The erosion rate has a loss of regularity at stagnation points, and this
inevitably leads to corners developing on the eroding bodies.
Therefore, as described in Section~\ref{sec:formulation} and our
previous work~\cite{qua-moo2018}, we introduce a curvature penalization
term and a Gaussian smoothing step.  For all examples, we use the
smoothing parameters $\eps=15/256$ and $\sigma=10/256$.

The common characteristic of each of the experiments is near-contact
between the eroding bodies, outer walls, and streamlines. We use our
numerical methods to describe and visualize the following examples:
\begin{itemize}
  \item{\bf Single Body Close to a Wall}: We consider a single eroding
  body close to the outer wall at $y=-1$.  We impose a shear flow
  centered at $y=-1$, we compare the eroding body's shape to a similar
  experiment of Spagnolie and Mitchell~\cite{mit-spa2017}.

@@

  \item{\bf 100 Bodies at a Medium Porosity}: We consider 100 eroding
  bodies with a medium porosity.  We compute the tortuosity, anomalous
  dispersion rates, and the pore throat size distributions.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a single eroding body close to a solid wall with the
shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
imposed on the outer wall. Mitchell and Spagnolie~\mit{spa-spa2017} did
a similar three-dimensional experiment using a second-order quadrature
method. Here, they initialized the eroding body so that its center $1.5$
times the body's radius above the solid wall.  In two dimensions, the
trapezoid rule is spectrally accurate, and an analogous initial
configuration is possible. However, unless the number of discretization
points is increased, the trapezoid rule becomes obsolete if the eroding
body is initialized closer to the solid wall. Therefore, to simulate
erosion near a solid wall at reasonable levels of discretization, we use
our hybrid method that combines the trapezoid rule and the Barycentric
quadrature rule.

We consider an eroding body that is initially a circle of radius
$r=0.4$.  We discretize the eroding body with $N_\iin=256$
discretization points, and its center is chosen so that the distance
between the body and the solid wall is $h/2$, $h/4$, and $h/10$, where
$h = 2\pi r/N_\iin$.  If we used the trapezoid rule and required an
error that is comparable to the Barycentric quadrature rule, the body
with an initial distance of $h/10$ from the solid wall would require
$6,400$ discretization points, and the outer wall would require $50,000$
discretization points.  For all three simulations, we use the time step
size $\Delta t=10^{-4}$.

In Figure~\ref{fig:NearWall}, we superimpose the eroding body's shape at
equispaced time steps.  For the three initial configurations, the shear
stress is positive for all time, but varies over several orders of
magnitude.  Therefore, we color the eroding body's boundary with the
logarithm of the shear stress.  Since the shear stress is always
positive, the erosion rate is smooth and corners do not develop, but
interesting shapes form nonetheless.  For instance, in the top half of
the body, there is a sudden increase in the shear stress, and this leads
to a region of high curvature.  This behavior is also present in three
dimensions~\cite{mit-spa2017} (Figure 7(c)).  The biggest difference
between the two- and three-dimensional results is the presence of a
recirculation zone.  In three dimensions, there is no recirculation
between the solid wall and the spherical body~\cite{cha-feu2003}, but
this is possible in two dimensions~\cite{chw-wu1975, hig1985}. To
visualize the flow, we plot the vorticity of the final time step from
Figure~\ref{fig:NearWall} in Figure~\ref{fig:NearWall_vort}.  The
recirculation is present in the region where the vorticity is smallest,
but its size is very small.

\begin{figure}[htbp]
\begin{center}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r1h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d5h_shear}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.38\textwidth}
\includegraphics[width =\textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{}
\end{subfigure}
\caption{\label{fig:NearWall} A single body eroding in a Stokes flow.
The color is the logarithm of the shear stress. Therefore, erosion is
fastest in the red regions (upper half) and slowest in the blue regions
(lower half).  The body is initialized at three different distances from
the lower wall: (a) $h$, (b) $h/2$, and (c) $h/10$.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r1h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d5h_vort}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics[width = \textwidth]{./figs/1b_0d4r0d1h_vort}
\caption{}
\end{subfigure}
\caption{\label{fig:NearWall_vort} The vorticity of the fluid with a
single body eroding at time t=0.1. The initial distance from the body to
the solid wall are: (a) $h$, (b) $h/2$, and (c) $h/10$.}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points and a time step size of $\Delta t = 10^{-4}$.  The background
flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 2.  Snapshots of the bodies at four equispaced
times are shown in Figure~\ref{fig:Eroding20vort}.  The color is the
vorticity, and the erosion rate is proportional to the vorticity.
Initially, several of the eroding bodies are closer to the outer wall
than the $5h$ threshold required to guarantee that the trapezoid rule
achieves machine precision.  In particular, the distance between bodies
1, 6, 13, and 15 and the outer wall is $1.3h$, $2.9h$, $2.8h$, and
$1.3h$, respectively, where $h$ is the arclength spacing of the outer
wall $\Gamma$.  In addition, the distance between several pairs of
eroding bodies, including 1 \& 6, 3 \& 9, 6 \& 8, and 14\& 18, is too
small to be accurately resolved with the trapezoid rule.  By using the
Barycentric quadrature rule, the interaction between these
nearly-touching bodies can be accurately resolved at the prescribed
resolution, and erosion can be simulated until all the bodies have
vanished.

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.225\textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} The erosion of 20 nearly touching
bodies, each discretized with $N_\iin=256$ points, in a Hagen-Poiseuille
flow.  The four snapshots are evenly spaced in time, and the color is
the fluid vorticity. In the fourth frame, bodies 4, 12, and 17 have
vanished completely, and body 19 has almost completely eroded.}
\end{center}
\end{figure}

Similar to our previous work, erosion causes the space between nearly
touching bodies to quickly expand, and flat faces develop along the
regions of near contact.  This qualitative behavior is present in
Figure~\ref{fig:Eroding20vort} between bodies 3 \& 4, 15 \& 16, and
others.  However, by resolving the interaction between bodies that are
much closer together than in our previous work, we observe that, at
least initially, very little erosion occurs between certain pairs of
bodies.  For instance the opening between bodies 1 \& 6, 3 \& 9, and 5
\& 13 grow much slower than the opening between bodies 15 \& 16.  A
common feature of the pores that grow slowly is that they are nearly
perpendicular to the main flow direction, and this results in a small
erosion rate.

Following our previous work~\cite{qua-moo2018}, we begin by analyzing
the effect of erosion on the area fraction and the flow rate.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  The general trend of the area fraction
resembles our previous work~\cite{qua-moo2018} (Figure 10(a)), but with
a larger initial area fraction.  In
Figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\cite{qua-moo2018} (Figure 10(b)), except that the initial flow
rate is an order of magnitude smaller because of the larger initial area
fraction.  Starting around normalized time $0.2$,
Figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as a power law.  Using a line of best
fit, the flow rate is approximately 
\begin{align} 
  U \approx \left(\frac{t}{t_f}\right)^{3.43},
\end{align}
which is the dashed line in Figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/flow_rate20dense}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The area fraction of 20
nearly touching eroding bodies versus normalized time. (b) The flow
rate, $U$, for a fixed pressure drop across the channel versus
normalized time.  The flow rate is initially very small, but it
eventually increases as a power law (dashed line) towards the flow rate
$U=1$ that occurs once all the bodies have eroded.}
\end{figure}

In Figure~\ref{fig:Eroding20vort}, we observe that erosion creates a
network of channels from the inlet to outlet where the velocity and
vorticity, and therefore erosion rate, are much larger than other
regions.  These channels can be further visualized by considering the
streamlines.  In Figure~\ref{fig:Eroding20tracer}, we freeze the
geometry at the second snapshot from Figure~\ref{fig:Eroding20vort}, and
we plot 200 streamlines that are initialized at $x=-1$ and equally
spaced in $y \in (-1,1)$.  The streamlines are shown at five different
times, and the final plot is a zoom in of the lower right quadrant of
the fifth time step.  Since we use a high-order quadrature rule and time
stepping method, we resolve streamlines that come very close to the
eroded bodies.  There are three clear regions where the streamlines are
concentrated, and this corresponds to regions of high velocity.  Two of
these regions are located between the bodies and the solid walls at
$y=\pm 1$, and the third cuts through the porous region with the upper
part of the channel formed by bodies 1, 4, 6, and 8.  Since the flow is
fastest in these regions, the shearing is largest, and this causes the
channels to continue to open fastest which can be observed in
Figure~\ref{fig:Eroding20vort}.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from Figure~\ref{fig:Eroding20vort}. The streamlines are
initialized at $x=-1$ and are equally spaced in $y \in (-1,1)$. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot.  Since we use the Barycentric
quadrature rule and the fourth-order Runge-Kutta method, some
streamlines are very close to the grains without penetrating them.}
\end{center}
\end{figure}

Next, we use the streamlines to compute the local tortuosity and the
anomalous dispersion rate of eroding geometries.  To achieve statistical
convergence, we use $N_p = 1000$ streamlines~\cite{bel-sal-rin1992}. To
compute the tortuosity, we require the velocity at the inlet $x=-1$.
These normalized velocities are plotted in
Figure~\ref{fig:Eroding20tort}(a) for the eroded geometry with porosity
62.9\% (see Figure~\ref{fig:Eroding20tort}(c)).  The velocities are
similar to the work of Matyka et al.~\cite{mat-kha-koz2008} (see Figure
4(a)), except that our cross-section, by construction, does not cut
through any of the grains.  Next, in Figure~\ref{fig:Eroding20tort}(b),
we plot the local tortuosity~\eqref{eqn:localTort} by calculating the
relative length of each streamline as it traverses the channel from
$x=-1$ to $x=1$.  The local tortuosity ranges from 1 to 1.27, meaning
that one of the streamlines is 27\% longer than it would have been if
the grains were absent.  The average streamline length is 9.79\% longer
than if the grains had been absent, so the tortuosity of the eroded
channel is $1.098$.  Again, comparing the local tortuosity to
Matyka~\cite{mat-kha-koz2008} (Figure 4(b)), the results are
qualitatively similar. However, since our initial cross-section does not
cut through the grains, the local tortuosity does not have any gaps.
The discontinuities in the local tortuosity happens when nearby
streamlines diverge on opposite sides of a grain.  In
Figure~\ref{fig:Eroding20tort}(c), we plot the pairs of streamlines that
correspond to the ten largest jumps in the local tortuosity.
Corresponding pairs are plotted in the same color.
